---
title: "Lab 7"
author: "Anne-Marie Parkinson"
date: "February 20, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = F,
                      warning = F)
```

# load packages
```{r}

#general
library(tidyverse)
library(here)
library(janitor)
library(plotly)

#spatial maps
library(tmap)
library(maptools)
library(sf)
library(sp)
library(spatstat)
library(raster)

#cluster analysis
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```


# load data
```{r}

#species data
voles <- read_sf(dsn = here("data","redtreevoledata"), 
                 layer = "ds033") %>% 
  dplyr::select(COUNTY) %>% 
  filter(COUNTY == "HUM") %>% 
  st_transform(crs = 4326)

st_crs(voles) # check projection
# hUmbolt county layer
humboldt <- read_sf(dsn = here("data", "redtreevoledata"), layer = "california_county_shape_file", crs = 4326) %>% 
  filter(NAME == "Humboldt") %>% 
  dplyr::select(NAME)
  
st_crs(humboldt) # check projection
  
# used two different crs methods. Depends on whether there is a crs assigned in the original dat or not, but the latter method works for either situation. 

```

```{r}
#initial visualization of the data
plot(voles)
plot(humboldt)

# tmap
tm_shape(humboldt) +
  tm_fill() +
  tm_shape(voles) +
  tm_dots(size = 0.2)

ggplot() +
  geom_sf(data = humboldt) +
  geom_sf(data = voles)

```

#convert vole events and humboldt polygon to point pattern + window

```{r}
voles_sp <- as(voles, "Spatial") #ensure R recognizes the voles data set as spatial data. Some functions dont like sf data

#voles_ppp <- as(voles_sp, "ppp") # errors with package, allison will update later

```

# cluser analysis

## K means

```{r}
#clean data
iris_nice <- iris %>% 
  clean_names()

# explore data
ggplot(iris_nice) +
  geom_point(aes(x = petal_length, y = petal_width, color = species))

# we think it looks like there are 3 clusters
# ask R how many clusters it thinks there are
number_est <- NbClust(iris_nice[1:4], # the brackets tells R to only conisder data in columns 1-4
                      min.nc = 2,
                      max.nc = 10,
                      method = "kmeans") # use ?NbClust to see all the diff methods that can be used


```


graph output = count of how many algorithms thought the data had x number of clusters. Here the algorithms think 2 groups is most likley, but R is NO substitute for assessing the data visually yourself and we think that sicne there are 3 species 3 clusters make sense


```{r}
# run kmeans
iris_km <- kmeans(iris_nice[1:4], 3)

# use iris_km$cluster to determine what cluster R thinks each observation belongs to
```
size outputs have to add up to 100

```{r}
# bind the cluster number together with the original data

iris_cl <- data.frame(iris_nice, cluster_no = factor(iris_km$cluster)) # can also use as.factor

# plot 
ggplot(iris_cl) +
  geom_point(aes(x = sepal_length, y = sepal_width, color = cluster_no))

```

```{r}
plot_ly(x = iris_cl$petal_length,
        y = iris_cl$petal_width,
        z = iris_cl$sepal_width,
        color = iris_cl$cluster_no,
        type = "scatter3d"
        )
#allisons code
plot_ly(x = iris_cl$petal_length, 
        y = iris_cl$petal_width, 
        z = iris_cl$sepal_width, 
        type = "scatter3d", 
        color = iris_cl$cluster_no, 
        symbol = ~iris_cl$species,
        marker = list(size = 3),
        colors = "Set1")
```

## Hierarchical cluster analysis

Relevant functions:

stats::hclust() - agglomerative hierarchical clustering
cluster::diana() - divisive hierarchical clustering

```{r}

# load data
wb_env <- read_csv(here("data", "wb_env.csv")) 

#sub-data: only keep top 20 emitters
wb_ghg_20 <- wb_env %>% 
  arrange(-ghg) %>% 
  head(20)

# scale data so all variables are on the same scale
wb_scaled <- as.data.frame(scale(wb_ghg_20[3:7]))

# Update data frame to add rownames (country name)
rownames(wb_scaled) <- wb_ghg_20$name

```

### find distances(create dissimilaraity matrix)
```{r}
#calc euclidean
diss <- dist(wb_scaled, method = "euclidean")

# use euclidean distances to do some complete agglomerate clustering
hc_complete <- hclust(diss, method = "complete")

# plot outputs
plot(hc_complete)

# nicer plot
ggdendrogram(hc_complete, rotate = T)

```
things clustered closer on the dendeogram = closer relation in multivariate space, countries further apart = more dissimilar


















